Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-04-29 paper/ms.tex
update ms

\begin{enumerate}
  \item Two-sided posteriors available for both $t_{\rm gyro}$ and
    $t_{\rm Li}$: then if their values were consistent within
    2$\sigma$, we judged them to be consistent; if they were
    consistent within 2-3$\sigma$, we suggest them to be ``maybe''
    consistent.  Otherwise, they are in disagreement.
\item If $t_{\rm Li}$ is a lower limit, we compare the 3$\sigma$
  $t_{\rm Li}$ lower limit with the $1\sigma$ upper limit from $t_{\rm
    gyro}$; if the two overlapped, we judge the age estimates to be
    consistent.
\item If $t_{\rm Li}$ provides a two-sided posterior, and no rotation
  is reported, we judge the age estimates to be inconsistent.
  This is because such constraints can only be provided for G and K
  stars $\lesssim$2\,Gyr old, and Kepler should have been sensitive to
  the rotation periods of such stars.
\end{enumerate}

In the entire sample, this yielded 
\allagesyesconsistent\ consistent cases,
\allagesmaybeconsistent\ ``maybe'' cases, and
\allagesnoconsistent\ inconsistent.
Restricting our attention to the $\lesssim$1\,Gyr stars, for which
lithium ages are more constraining,
the relevant numbers are
\minageltonegyryesconsistent, \minageltonegyrmaybeconsistent, and
\minageltonegyrnoconsistent.
Rephrased, in the overall sample, \fracconsistentallages\%
of stars have consistent rotation and lithium-based ages
(\fracpotentiallyconsistentallages\% potentially consistent).
In the sample of sub-gigayear stars,
\fracconsistentminageltonegyr\ of these ages are
consistent (\fracpotentiallyconsistentminageltonegyr\% potentially
consistent).
We feel it necessary to distinguish ``maybe'' consistent cases because
of lingering uncertainties in how the dispersion of the lithium data
evolve over the first gigayear.
In Appendix~\ref{app:inconsistent}, we discuss the
\minageltonegyrnoconsistent\ discrepant cases in detail.

>>>>

\begin{enumerate}
  \item Two-sided posteriors available for both $t_{\rm gyro}$ and
    $t_{\rm Li}$: if their values were consistent within
    2$\sigma$, we judged them to be consistent; if they were
    consistent within 2-3$\sigma$, we suggest them to be ``maybe''
    consistent.  Otherwise, they are in disagreement.
\item If $t_{\rm Li}$ is a lower limit, we compare the
  $t_{\rm Li}$ lower limit with the $1\sigma$ upper limit from $t_{\rm
    gyro}$; if the two overlapped, we judge the age estimates to be
    consistent.
\item If $t_{\rm Li}$ provides a two-sided posterior, and no rotation
  is reported, we judge the age estimates to be inconsistent.
  This is because this lithium constraint can only be provided for G and K
  stars $\lesssim$2\,Gyr old, and Kepler should have been sensitive to
  the rotation periods of such stars.
\end{enumerate}

In the entire sample, this yielded 
\allagesyesconsistent\ consistent cases,
\allagesmaybeconsistent\ ``maybe'' cases, and
\allagesnoconsistent\ inconsistent.
Restricting our attention to the $\lesssim$1\,Gyr stars, for which
lithium ages are more constraining,
the respective counts are
\minageltonegyryesconsistent, \minageltonegyrmaybeconsistent, and
\minageltonegyrnoconsistent.
Rephrased, in the overall sample, \fracconsistentallages\%
of stars have consistent rotation and lithium-based ages
(\fracpotentiallyconsistentallages\% potentially consistent).
In the sample of sub-gigayear stars,
\fracconsistentminageltonegyr\% of the ages are
consistent (\fracpotentiallyconsistentminageltonegyr\% potentially
consistent).
We feel it necessary to distinguish ``maybe'' consistent cases because
of lingering uncertainties in how the dispersion of the lithium data
evolve over the first gigayear.
In Appendix~\ref{app:inconsistent}, we discuss the
\minageltonegyrnoconsistent\ discrepant sub-gigayear stars in detail.
